Pillar 1 - Sequential Pattern Mining

### 2019-01-30 Pillar 1 - Sequential Pattern Mining.Rmd
Last touch sequence rules: saves results, removes unnecessary objects and old code

#
# Generating asocciation rules
AR_taken <- my_apriori(transactions)

@@

AR_not_taken <- my_apriori(transactions_Taken)

#
# Save association rules
save(AR_taken, AR_PF, AR_HL, AR_not_taken,
     file = "App/ARR.RDATA")

@@

### CSPADE Algorithm
We apply cspade algorithm
```{r}
# Transform rules from ruleInduction into a readable data frame
clean_IR <- function(rules){
  rules %>%
    as("data.frame") %>%
    # filter(
    #   count >= 1
    # ) %>%
    mutate(
      rule = str_remove_all(rule, pattern = "[<]"),
      rule = str_remove_all(rule, pattern = "[{]"),
      rule = str_remove_all(rule, pattern = "[}]"),
      rule = str_remove_all(rule, pattern = "[>]"),
      `rhs.support` = confidence / lift
      ) %>%
    separate(rule, into = c("lhs", "rhs"), sep="=") %>%
    mutate_at(
      c("support", #"lhs.support",
        "confidence", "rhs.support", "lift"),
      funs(round(., 5))
    ) %>%
    select(
      lhs, rhs,
      support, 
      #count, lhs.support, #figure out how to get these
      confidence, rhs.support,
      lift
    ) #%>%
    #arrange(
    #  desc(count)
    #)

}
```
We apply cspade followed by rule induction
```{r}
# Run the cspade algorithm with desired parameters. Then run the ruleInduction algorithm with desired parameters.
my_IR <- function(data){
  data %>%
  cspade(
    parameter = list(
      support = 0.01, # min support of a sequence
      maxsize = 1, # max number of items of an element of a sequence
      maxlen  = 2, # max number of element of a sequence
      mingap  = 1, # min time difference between consecutive element of a sequence
      maxgap  = 1e4 # max time difference between consecutive element of a sequence
      #maxwin  = 1e4  # max time difference between any two elements of a sequence # WARNING: 'maxwin' disabled
      ),
    control = list(
      verbose = FALSE
      )
    ) %>%
    ruleInduction(
      confidence = 0, #we need to play with the parameters.
      control    = list(verbose = FALSE)
      ) %>%
    clean_IR
}

#Generating induced rules
IR    <- my_IR(data = sequences   )
IR_PF <- my_IR(data = sequences_PF)
IR_HL <- my_IR(data = sequences_HL)

#Saving induced rules:
save(
  IR = list(
    IR = IR,
    IR_PF = IR_PF,
    IR_HL = IR_HL
    ),
  file = "Output/IR.RDATA")

rm(clean_IR, my_IR,
   sequences, sequences_PF, sequences_HL,
   IR, IR_PF, IR_HL)

@@

##calculate appropriate measures:
```{r}
n_students <- length(unique(d_sequence_transactions$`Student ID`))

sequence_rules <- pass_rules_df %>%
  separate(LHS, into=c("LHS Course ID", "LHS Score"), sep="_", remove= FALSE) %>%
  separate(RHS, into=c("RHS Course ID", "RHS Score"), sep="_", remove= FALSE) %>%
  mutate(`RHS Course ID`= substr(`RHS Course ID`, start=4, stop=10))%>%
  left_join(lookup_table, by = c("RHS Course ID"="Course ID")) %>%

  # Compute statistics
  group_by(LHS,`RHS Course ID`) %>%
  mutate(Prob_both = sum(support), 
         Conf_corr = support/Prob_both,
         n_rule = support * n_students,
         n_both =Prob_both * n_students,
         lift_diff= Conf_corr - Prop_low,
         lift_corr=Conf_corr / Prop_low) %>%
  ungroup() %>%

  # Filter and edit output
  filter(str_detect(`RHS Score`, pattern = "low"),
         str_detect(`LHS Score`, pattern = "low"),
         n_rule >= 5,
         Conf_corr > 0.2) %>%
  select(LHS, RHS, support, n_rule, Prob_both, n_both, Prop_low, Conf_corr, lift_corr, lift_diff) %>%
  mutate(support = round(support,3),
         Prob_both =round(Prob_both,3),
         Prop_low =round(Prop_low,3),
         Conf_corr= round(Conf_corr,3),
         lift_diff= round(lift_diff,3),
         lift_corr= round(lift_corr,3)) %>%
  arrange(lift_diff)

```
##Save data
```{r}
save(sequence_rules,
     file = "sequence_rules.RDATA")
```

##Arrange data for plots:
```{r}
for_plots <-  manual_confidence_rules %>%
  group_by(LHS)%>%
  mutate(LHS_count=n())
```
#Visualizing rules
```{r}
#look at counts
ggplot(for_plots, aes(LHS))+
  geom_bar()

@@

##TO MOVE UP (These are ready:)
Rounds grades and add new row from original grade up to 10.
NOTE: includes mandatory courses.
```{r}
distributed_transcript <- d_transcript %>%
  filter(!is.na(Grade), Grade>=0) %>%
  mutate(Round_grade = paste("g", round(Grade,0),sep=""),
         Values = T) %>%
  spread(key = "Round_grade", value = Values) %>%
  mutate(g1  = g0|g1,
         g2  = g1|g2,
         g3  = g2|g3,
         g4  = g3|g4,
         g5  = g4|g5,
         g6  = g5|g6,
         g7  = g6|g7,
         g8  = g7|g8,
         g9  = g8|g9,
         g10 = g9|g10,
         ) %>%
  gather(key="Gr_string", value="Gr_bool", g0, g1, g2, g3, g4,g5,g6,g7,g8,g9,g10, na.rm=T) %>%
  select(-Gr_bool )%>%
  separate(Gr_string, into=c("Trash", "Grade_new"), sep="g")%>%
  select(-Trash) %>%
  mutate(Grade= as.numeric(Grade_new))%>%
  select(-Grade_new) #%>%
  #to check that the grades appear from original grade to 10.
  #arrange(`Student ID`, `Course ID`) 
```

##Not ready to move up:

>>>>

#
# Generating asocciation rules
AR <- my_apriori(transactions)

@@

AR_Taken <- my_apriori(transactions_Taken)

#
# Save association rules
AR <- list(
    AR = AR,
    AR_PF = AR_PF,
    AR_HL = AR_HL,
    AR_Taken = AR_Taken
    )

save(AR, file = "App/AR.RDATA")

@@

### CSPADE Algorithm
```{r}
# Transform rules from ruleInduction into a readable data frame
n_students <- length(unique(d_transactions$sequenceID))

clean_SR<- function(rules){

  rules %>%
    as("data.frame") %>%
    mutate(
      rule = str_remove_all(rule, pattern = "[<]"),
      rule = str_remove_all(rule, pattern = "[{]"),
      rule = str_remove_all(rule, pattern = "[}]"),
      rule = str_remove_all(rule, pattern = "[>]")
      ) %>%
    separate(rule, into = c("lhs", "rhs"), sep="=")  %>%
    select(
      lhs, rhs,
      support,
      confidence,
      lift
    ) %>%
    separate(lhs, into=c("LHS Course ID", "LHS Quality"), sep="_", remove= F) %>%
    separate(rhs, into=c("RHS Course ID", "RHS Quality"), sep="_", remove= F) %>%
    # Compute statistic
    group_by(lhs,`RHS Course ID`) %>%
    mutate(
      lhs.rhsCourse.support = sum(support),
      confidence.corr = support / lhs.rhsCourse.support,
      count = support * n_students,
      lhs.rhsCourse.count = lhs.rhsCourse.support * n_students

      # TODO: coñpute rhs.support
      # lift.corr = confidence.corr / rhs.support

  ) %>%
  ungroup() %>%
    mutate_at(
      c("support",
        "confidence", "lift"),
      funs(round(., 5))
    )

```
We apply cspade followed by rule induction
```{r}
# Run the cspade algorithm with desired parameters. Then run the ruleInduction algorithm with desired parameters.
my_SR<- function(data){
 data %>%
  cspade(
    parameter = list(
      support = 0, # min support of a sequence
      maxsize = 1, # max number of items of an element of a sequence
      maxlen  = 2, # max number of element of a sequence
      mingap  = 1, # min time difference between consecutive element of a sequence
      maxgap  = 1e4 # max time difference between consecutive element of a sequence
      #maxwin  = 1e4  # max time difference between any two elements of a sequence # WARNING: 'maxwin' disabled
      ),
    control = list(
      verbose = FALSE
      )
    ) %>%
    ruleInduction(
      confidence = 0, #we need to play with the parameters.
      control    = list(verbose = FALSE)
      ) %>%
    clean_SR
}

#Generating sequence rules
SR   <- my_SR(data = sequences   )
SR_PF <- my_SR(data = sequences_PF)
SR_HL <- my_SR(data = sequences_HL)

#Saving sequence rules
SR <- list(
    SR = SR,
    SR_PF = SR_PF,
    SR_HL = SR_HL
    )

save(SR, file = "App/SR.RDATA")

#clean unnecessary objects
rm(clean_SR, my_SR, 
   sequences, sequences_PF, sequences_HL, 
   n_students,
   SR, SR_PF, SR_HL)

@@

#Visualizing rules
```{r}
#look at counts
ggplot(for_plots, aes(LHS))+
  geom_bar()

@@

##TO MOVE UP (These are ready:)
Rounds grades and add new row from original grade up to 10.
NOTE: includes mandatory courses.
```{r}
distributed_transcript <- d_transcript %>%
  filter(!is.na(Grade), Grade>=0) %>%
  mutate(Round_grade = paste("g", round(Grade,0),sep=""),
         Values = T) %>%
  spread(key = "Round_grade", value = Values) %>%
  mutate(g1  = g0|g1,
         g2  = g1|g2,
         g3  = g2|g3,
         g4  = g3|g4,
         g5  = g4|g5,
         g6  = g5|g6,
         g7  = g6|g7,
         g8  = g7|g8,
         g9  = g8|g9,
         g10 = g9|g10
         ) %>%
  gather(key="Gr_string", value="Gr_bool", g0, g1, g2, g3, g4,g5,g6,g7,g8,g9,g10, na.rm=T) %>%
  select(-Gr_bool )%>%
  separate(Gr_string, into=c("Trash", "Grade_new"), sep="g")%>%
  select(-Trash) %>%
  mutate(Grade= as.numeric(Grade_new))%>%
  select(-Grade_new) #%>%
  #to check that the grades appear from original grade to 10.
  #arrange(`Student ID`, `Course ID`) 
```
